Handling String-Number Conversion in String Constraint Solving

### 2019-11-08 draft.tex
before removoving epsilons from product constr

\newcommand{\todo}[1]{{\color{blue}TODO: #1}}
\newcommand{\lh}[1]{{\color{orange}Lukas: #1}}
\newcommand{\sti}[1]{\mbox{\textsf{toInt}($#1$)}}
\newcommand{\its}[1]{\mbox{\textsf{toStr}($#1$)}}
\newcommand{\varn}{\mbox{$\mathbb{V}_{\mathbb{Z}}$}}
\newcommand{\vars}{\mbox{$\mathbb{V}_{\Sigma^*}$}}
\newcommand{\cvars}{\mbox{$\mathbb{V}_{\Sigma_\epsilon}$}}
\newcommand{\pvars}{\mbox{$\mathbb{V}_{\sharp}$}}
\newcommand{\cvar}{\mbox{$V_{\Sigma_\epsilon}$}}
\newcommand{\cvarone}{\mbox{$V^1_{\Sigma_\epsilon}$}}
\newcommand{\cvartwo}{\mbox{$V^2_{\Sigma_\epsilon}$}}
\newcommand{\modelsof}[1]{[\![#1]\!]}
\newcommand{\true}{\mbox{$\mathsf{true}$}}
\newcommand{\false}{\mbox{$\mathsf{false}$}}
\newcommand{\enc}[1]{[\![#1]\!]}
%\newcommand{\parikhof}[1]{\mathit{Parikh}{(#1)}}
\newcommand{\parikhof}[1]{\mathbb{P}{(#1)}}
\newcommand{\parikhwof}[2]{|#1|_{#2}}
%\newcommand{\semof}[1]{||#1||}
\newcommand{\semof}[1]{\modelsof{#1}}
\newcommand{\decode}{\mathit{decode}}
\newcommand{\pformulaof}[1]{\Phi_{#1}}
\newcommand{\pim}{\alpha}
\newcommand{\syncof}[2]{A \Cap B}
\newcommand{\nat}{\mathbb{N}}

@@

Since every string $w \in \semof{A}$ is obtained from a parametric string $x\in L(A)$ and an interpretation of character varriables $I:\cvar\rightarrow\Sigma_\epsilon$ such that $w = I(x)$, and since $x = \decode(\parikhof x)$, 
we may say that the united assignment $I \cup \parikhof x$ encodes $w$. Lemma~\ref{lemma:parikh} also implies that two different strings are encoded by different united assignments, and hence we may extend the decoding function $\decode$ to united assignments so that $\decode(I \cup x) = w$.
Encodings of strings in $\semof{A}$ are encodings of strings of $L(A)$ paired with arbitrary interpretations of $\cvar$, 
hence $\semof A$ is characterised by $\parikhof A$ interpreted over variables $\cvar\cup\pvars$: 
%\begin{lemma}\ref{lemma:decw}
%For a PFA $A$, $\semof{A} = \{\decode(I\cup \pim) \mid \pim\in\semof{\pformulaof A} \land I:\cvar\rightarrow \Sigma_\epsilon\}$.
%\end{lemma}
\begin{lemma}
For a PFA $A$, $\semof{A} = \{\decode(I\cup \pim) \mid I\cup\pim\in\semof{\pformulaof A}_{\cvar\cup\pvars}\}$.
\end{lemma}

\paragraph{Synchronisation of PFA}
The following construction of the so called synchronisation formula of two PFA $A$ and $B$ is a crucial component of bulding overapproximations of string constraints. 
Synchronisation formula is a formula $\syncof A B$ over $\pvars\cup\cvarone\cup \cvartwo$ encoding the semantic intersection $\semof A \cap \semof B$. Besides the intersection, it contains a more precise information about the synchronisation of $A$ and $B$, 
namely, it characterises the set of pairs of encodings $(I\cup \pim)\in \parikhof A$ and $(I'\cup \pim')\in\parikhof B$ such that 
$\decode(I\cup \pim) = \decode(I'\cup \pim')$ which constitute the intersection.

Assume that $A$ and $B$ are over disjoint sets of variables. 
We build so called asynchronous product machine of $A$ and $B$ and then extract the formula from its Parikh image.
%
The product machine uses $Q\times P$ as the set of states and $(\cvarone \cup \{\epsilon\}) \times (\cvartwo \cup \{\epsilon\} )$ as the alphabet. If a transition $((q_1,p_1), (v,v'),(q_2,p_2))$ is taken, it means the character variable $v$ and $v'$ should be assigned the same value and this value leads $S_n$ from state $q_1$ to state $q_2$ and $S_m$ from $p_1$ to $p_2$.

More precisely, the product machine is a tuple $M = (Q\times P, T^2, (\cvarone \cup \{\epsilon\}) \times (\cvartwo \cup \{\epsilon\} ), (q_i,p_i),(q_f,p_f))$, where the transition relation $T^2$ is the minimal set satisfying the following.

\begin{itemize}
\item If $(q_1,v,q_2) \in T$, then for all states $p\in P$, we have $((q_1,p),(v,\epsilon),(q_2,p))\in T^2$.
\item If $(p_1,v,p_2) \in T'$, then for all states $q\in Q,((q,p_1)$, we have $(\epsilon,v),(q,p_2))\in T^2$.
\item If $(q_1,v,q_2) \in T$ and $(p_1,v',p_2) \in T'$, then we have $((q_1,p_1),(v,v'),(q_2,p_2))\in T^2$.
\end{itemize}	
The first two cases correspond to the situation when the character variable $v$ is assigned $\epsilon$. In such cases, only one automaton changes its state.
%
We compute a Presburger formula $\pformulaof M$ over the set of variables $\{\#(v,v')\mid (v,v') \in \cvarone \times \cvartwo\}$ of the product machine. 
%We use $\#v$ and $\#v'$ to denote the number of occurrences of $v_i$ and $v_i^{(j)}$, respectively. 
The following formulae establish the relation of the number of occurrences of symbols between the SFAs and the product machine.
\begin{enumerate}
	\item for all $v \in \cvarone$, $\#(v) = \sum_{v' \in {\cvartwo \cup \{\epsilon\}}} \#{(v,v')}$   
	\item for all $v \in \cvartwo$, $\#(v) = \sum_{v' \in {\cvarone \cup \{\epsilon\}}} \#{(v',v)}$ 
\end{enumerate}

@@

The correctness of this construction is expressed by the following lemma
\begin{lemma}
For $\pim:\pvars\rightarrow \nat$, $\pim':\pvars\rightarrow\nat$, $I:\cvarone\rightarrow\Sigma$, $I':\cvartwo\rightarrow\Sigma$, 
we have $(I\cup\pim)\cup(I'\cup\pim')\in\semof \Psi$ if and only if $\decode(I\cup\pim) = \decode(I'\cup\pim')$.
\end{lemma}
We actually have even something stronger, namely, that, $\Psi$ has all possible allignments. How do I say this ... or do I have to say this?

@@

Such product automaton uses $Q\times P$ as the set of states and $(\cvarone \cup \{\epsilon\}) \times (\cvartwo \cup \{\epsilon\} )$ as the alphabet. If a transition $((q_1,p_1), (v_i^{(j)},v_{i'}^{(j')}),(q_2,p_2))$ is taken, for some $i,j,i',j' \in \mathbb{N}$, it means the character variable $v_i$ and $v_{i'}$ (and so all of their indexed variants) should be assigned the same value and this value leads $S_n$ from state $q_1$ to state $q_2$ and $S_m$ from $p_1$ to $p_2$.

More precisely, the product automaton is a tuple $(Q\times P, T^2, (\cvarone \cup \{\epsilon\}) \times (\cvartwo \cup \{\epsilon\} ), (q_i,p_i),(q_f,p_f))$, where the transition relation $T^2$ is the minimal set satisfying the following.

\begin{itemize}
\item If $(q_1,v_i^{(j)},q_2) \in T$, then for all states $p\in P$, we have $((q_1,p),(v_i^{(j)},\epsilon),(q_2,p))\in T^2$.
\item If $(p_1,v_i^{(j)},p_2) \in T'$, then for all states $q\in Q,((q,p_1)$, we have $(\epsilon,v_i^{(j)}),(q,p_2))\in T^2$.
\item If $(q_1,v_i^{(j)},q_2) \in T$ and $(p_1,v_{i'}^{(j')},p_2) \in T'$, then we have $((q_1,p_1),(v_i^{(j)},v_{i'}^{(j')}),(q_2,p_2))\in T^2$.
\end{itemize}	
The first two cases correspond to the situation when the character variable $v_i^{(j)}$ is assigned $\epsilon$. In such cases, only one automaton changes its state.

>>>>

\newcommand{\nat}{\mathbb{N}}
\newcommand{\todo}[1]{{\color{blue}TODO: #1}}
\newcommand{\lh}[1]{{\color{orange}Lukas: #1}}
\newcommand{\sti}[1]{\mbox{\textsf{toInt}($#1$)}}
\newcommand{\its}[1]{\mbox{\textsf{toStr}($#1$)}}
\newcommand{\varn}{\mbox{$\mathbb{V}_{\mathbb{Z}}$}}
\newcommand{\vars}{\mbox{$\mathbb{V}_{\Sigma^*}$}}
\newcommand{\cvars}{\mbox{$\mathbb{V}_{\Sigma_\epsilon}$}}
\newcommand{\pvars}{\mbox{$\mathbb{V}_{\sharp}$}}
\newcommand{\cvar}{\mbox{$V_{\Sigma_\epsilon}$}}
%\newcommand{\cvarone}{\mbox{$V^1_{\Sigma_\epsilon}$}}
%\newcommand{\cvartwo}{\mbox{$V^2_{\Sigma_\epsilon}$}}
\newcommand{\cvarone}{\mbox{$V_{\Sigma_\epsilon}$}}
\newcommand{\cvartwo}{\mbox{$V'_{\Sigma_\epsilon}$}}
\newcommand{\modelsof}[1]{[\![#1]\!]}
\newcommand{\true}{\mbox{$\mathsf{true}$}}
\newcommand{\false}{\mbox{$\mathsf{false}$}}
\newcommand{\enc}[1]{[\![#1]\!]}
%\newcommand{\parikhof}[1]{\mathit{Parikh}{(#1)}}
\newcommand{\parikhof}[1]{\mathbb{P}{(#1)}}
\newcommand{\parikhwof}[2]{|#1|_{#2}}
%\newcommand{\semof}[1]{||#1||}
\newcommand{\semof}[1]{\modelsof{#1}}
\newcommand{\decode}{\mathit{decode}}
\newcommand{\pformulaof}[1]{\Phi_{#1}}
\newcommand{\pim}{\alpha}
\newcommand{\syncop}{\Cap}
\newcommand{\syncof}[2]{A \syncop B}
\newcommand{\syncfof}[2]{\Psi_{\syncof {#1} {#2}}}
\newcommand{\syncfof}[2]{\Psi_{\syncof {#1} {#2}}}
\newcommand{\syncT}{T_\syncop}

@@

Since every string $w \in \semof{A}$ is obtained from a parametric string $x\in L(A)$ and an interpretation of character varriables $I:\cvar\rightarrow\Sigma_\epsilon$ such that $w = I(x)$, and since $x = \decode(\parikhof x)$, 
we may say that the assignment $I \cup \parikhof x$ uniting the interpretation $I$ and the Parikh image of the parametric string $x$ encode $w$. Lemma~\ref{lemma:parikh} also implies that two different strings are encoded by different united assignments (the Parikh image or the interpretation must be different), and hence we may extend the decoding function $\decode$ to united assignments so that $\decode(I \cup x) = w$.
Encodings of strings in $\semof{A}$ are encodings of strings of $L(A)$ paired with arbitrary interpretations of $\cvar$, 
hence $\semof A$ is characterised by models of $\parikhof A$ interpreted over variables $\cvar\cup\pvars$ (variables of $\cvar$  may have arbitrary values since $\parikhof A$ does not constraint them): 
%\begin{lemma}\ref{lemma:decw}
%For a PFA $A$, $\semof{A} = \{\decode(I\cup \pim) \mid \pim\in\semof{\pformulaof A} \land I:\cvar\rightarrow \Sigma_\epsilon\}$.
%\end{lemma}
\begin{lemma}
For a PFA $A$, $\semof{A} = \{\decode(I\cup \pim) \mid I\cup\pim\in\semof{\pformulaof A}_{\cvar\cup\pvars}\}$.
\end{lemma}

\paragraph{Synchronisation of PFA}
The following construction of the so called synchronisation formula of two PFA $A$ and $B$ is a crucial component of bulding overapproximations of string constraints. 
Synchronisation formula is a formula $\syncfof A B$ over $\pvars\cup\cvarone\cup \cvartwo$ encoding the semantic intersection $\semof A \cap \semof B$. Besides the intersection, it contains a more precise information about the synchronisation of $A$ and $B$, 
namely, it characterises the set of pairs of encodings $(I\cup \pim)\in \parikhof A$ and $(I'\cup \pim')\in\parikhof B$ such that 
$\decode(I\cup \pim) = \decode(I'\cup \pim')$ which constitute the intersection.

Assume that $A = (Q,T,\cvarone,q_i,q_f)$ and $B = (P,T',\cvartwo,p_i,p_f)$ are two SFA over disjoint sets of variables, $\cvarone \cap \cvartwo = \emptyset$. 
We build so called asynchronous product automaton $\syncof A B$ of $A$ and $B$ and then extract the formula $\syncfof A B$ from its Parikh image.
%
The product automaton uses $Q\times P$ as the set of states and $(\cvarone \cup \{\epsilon\}) \times (\cvartwo \cup \{\epsilon\} )$ as the alphabet. If a transition $((q_1,p_1), (v,v'),(q_2,p_2))$ is taken, it means the character variable $v$ and $v'$ should be assigned the same value and this value leads $A$ from state $q_1$ to state $q_2$ and $B$ from $p_1$ to $p_2$.

More precisely, the product automaton is a tuple $\syncof A B = (Q\times P, \syncT, (\cvarone \cup \{\epsilon\}) \times (\cvartwo \cup \{\epsilon\} ), (q_i,p_i),(q_f,p_f))$, where the transition relation $\syncT$ is the minimal set satisfying the following.

\begin{itemize}
\item If $(q_1,v,q_2) \in T$, then for all states $p\in P$, we have $((q_1,p),(v,\epsilon),(q_2,p))\in \syncT$.
\item If $(p_1,v,p_2) \in T'$, then for all states $q\in Q,((q,p_1)$, we have $(\epsilon,v),(q,p_2))\in \syncT$.
\item If $(q_1,v,q_2) \in T$ and $(p_1,v',p_2) \in T'$, then we have $((q_1,p_1),(v,v'),(q_2,p_2))\in \syncT$.
\end{itemize}	
The first two cases correspond to the situation when the character variable $v$ is assigned $\epsilon$. In such cases, only one automaton changes its state.
%
We compute a Presburger formula $\pformulaof {\syncof A B}$ over the set of variables $\{\#(v,v')\mid (v,v') \in \cvarone \times \cvartwo\}$ of the product automaton. 
%We use $\#v$ and $\#v'$ to denote the number of occurrences of $v_i$ and $v_i^{(j)}$, respectively. 
The following formulae establish the relation of the number of occurrences of symbols between the SFAs and the product automaton.
\begin{enumerate}
	\item for all $v \in \cvarone$, $\#(v) = \sum_{v' \in {\cvartwo \cup \{\epsilon\}}} \#{(v,v')}$   
	\item for all $v \in \cvartwo$, $\#(v) = \sum_{v' \in {\cvarone \cup \{\epsilon\}}} \#{(v',v)}$ 
\end{enumerate}

@@

The correctness of this construction is expressed by the following lemma
\begin{lemma}
For $\pim:\pvars\rightarrow \nat$, $\pim':\pvars\rightarrow\nat$, $I:\cvarone\rightarrow\Sigma$, $I':\cvartwo\rightarrow\Sigma$, 
we have $(I\cup\pim)\cup(I'\cup\pim')\in\semof \syncofof A B$ if and only if $\decode(I\cup\pim) = \decode(I'\cup\pim')$.
\end{lemma}
We actually have even something stronger, namely, that, $\syncofof A B$ has all possible allignments. How do I say this ... or do I have to say this?

@@

Such product automaton uses $Q\times P$ as the set of states and 
§$(\cvarone \cup \{\epsilon\}) \times (\cvartwo \cup \{\epsilon\} )$ as the alphabet. If a transition $((q_1,p_1), (v_i^{(j)},v_{i'}^{(j')}),(q_2,p_2))$ is taken, for some $i,j,i',j' \in \mathbb{N}$, it means the character variable $v_i$ and $v_{i'}$ (and so all of their indexed variants) should be assigned the same value and this value leads $A$ from state $q_1$ to state $q_2$ and $B$ from $p_1$ to $p_2$.

More precisely, the product automaton is a tuple $\syncof A B = (Q\times P, \syncT, (\cvarone \cup \{\epsilon\}) \times (\cvartwo \cup \{\epsilon\} ), (q_i,p_i),(q_f,p_f))$, where the transition relation $\syncT$ is the minimal set satisfying the following.

\begin{itemize}
\item If $(q_1,v_i^{(j)},q_2) \in T$, then for all states $p\in P$, we have $((q_1,p),(v_i^{(j)},\epsilon),(q_2,p))\in \syncT$.
\item If $(p_1,v_i^{(j)},p_2) \in T'$, then for all states $q\in Q,((q,p_1)$, we have $(\epsilon,v_i^{(j)}),(q,p_2))\in \syncT$.
\item If $(q_1,v_i^{(j)},q_2) \in T$ and $(p_1,v_{i'}^{(j')},p_2) \in T'$, then we have $((q_1,p_1),(v_i^{(j)},v_{i'}^{(j')}),(q_2,p_2))\in \syncT$.
\end{itemize}	
The first two cases correspond to the situation when the character variable $v_i^{(j)}$ is assigned $\epsilon$. In such cases, only one automaton changes its state.
